
\documentclass{ltxdoc}

% Copyright 2003, 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% This program can be redistributed and/or modified under the terms
% of the GNU Public License, version 2.

\def\version{2.99}
\def\pgfversion{0.63}
\def\xcolorversion{2.00}

\usepackage{pgf,xcolor}
\usepackage[left=2.25cm,right=2.25cm,top=2.5cm,bottom=2.5cm,nohead]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[pdfborder={0 0 0},bookmarksnumbered]{hyperref}
\usepackage[latin1]{inputenc}
\usepackage{pifont}
\usepackage{makeidx}

\input{beamerug-macros}

\makeindex

%\includeonly{beamerug-introduction}

\begin{document}

\index{Themes|see{Presentation themes}}
\index{Templates|see{Beamer templates}}
\index{Colors|see{Beamer colors}}
\index{Fonts|see{Beamer fonts}}
\index{Beamer elements|see{Beamer templates, colors, and fonts}}
\index{Elements|see{Beamer templates, colors, and fonts}}
\index{Template inserts|see{Inserts}}

\title{User's Guide to the Beamer Class, Version \version\\
\Large\href{http://latex-beamer.sourceforge.net}{\texttt{http://latex-beamer.sourceforge.net}}}
\author{Till Tantau\\
  \href{mailto:tantau@users.sourceforge.net}{\texttt{tantau@users.sourceforge.net}}}

\maketitle

\tableofcontents


\include{beamerug-introduction}



\part{Getting Started}

This part helps you getting started. It starts with an explanation of
how to install the class. Hopefully, this will be very simple, with a
bit of luck the whole class is already correctly installed! You will
also find an explanation of special things you should consider when
using certain other packages.

Next, a short tutorial is given that explains most of the features
that you'll need in a typical presentation. Following the tutorial you
will find a ``possible workflow'' for creating a
presentation. Following this workflow may help you avoid problems
later on.

This part includes a guidelines sections. Following these guidelines
can help you create good presentations (no guarantees, though). This
guideline section is kept as general as possible; most what is said in
that section applies to presentations in general, independently of
whether they have been created using \beamer\ or not.

At the end of this part you will find a summary of the solutions
templates that come with \beamer. You can use solutions templates to
kick-start the creation of your presentation.

\include{beamerug-installation}
\include{beamerug-tutorial}
\include{beamerug-workflow}
\include{beamerug-guidelines}
\include{beamerug-solutions}



\part{Building a Presentation}

This part contains an explanation of all the commands that are used to
create presetations. It starts with a section treating the commands
and environments used to create \emph{frames}, the basic building
blocks of presentations. Next, the creation of overlays is
explained.

The following three sections concern commands and methods
of \emph{structuring} a presentation. In order, the \emph{static
  global} structure, the \emph{interactive global} structure, and the
\emph{local} structure are treated.

Two further sections treat graphics and animations. Much of the
material in these sections applies to other packages as well, not just
to \beamer.

At the end of this part, ``nonpresentation material'' is treated. This
includes handouts or lecture notes. Several commands can be used to
facilitate the creation of such handouts or lecture notes.

\include{beamerug-frames}
\include{beamerug-overlays}
\include{beamerug-globalstructure}
\include{beamerug-interaction}
\include{beamerug-localstructure}
\include{beamerug-graphics}
\include{beamerug-animations}
\include{beamerug-nonpresentation}



\part{Changing the Way Things Look}

\beamer\ offers ways to change the appearance of an appearance at all
levels of detail. On the top level, \emph{themes} can be used to
globally change the appearance conveniently. On the bottom level,
\emph{templates} allow you to specify the appearance of every miniute
detail individually.

To important aspects of the ``appearance'' of a presentation are
treated in two extra sections: colors and fonts. Here, too, color and
font themes can be used to globally change the colors or fonts used in
a presentation, while you can still change the color or font of, say,
block titles independently of everything else. 

\include{beamerug-themes}
\include{beamerug-elements}
\include{beamerug-color}
\include{beamerug-fonts}



\part{Howtos}

This part contains explanations-of-how-to-do-things (commonly known as
\emph{howtos}). These explanations are not really part of the
``\beamer-core.'' Rather, they explain how to use \beamer\ to achieve
a certain effect or how get something special done.

The first howto is about tricky uncovering situtations.

The second howto explains how you can import (parts or) presentations
created using some other \LaTeX-presentation class, like \prosper. 

\include{beamerug-tricks}
\include{beamerug-emulation}


\printindex

\end{document}
