% Copyright 2003--2007 by Till Tantau
% Copyright 2010 by Vedran Mileti\'c
% Copyright 2012,2013,2015 by Vedran Mileti\'c, Joseph Wright
% Copyright 2016 by Joseph Wright
% Copyright 2018 by Louis Stuart, Joseph Wright
%
% This file may be distributed and/or modified
%
% 1. under the LaTeX Project Public License and/or
% 2. under the GNU Free Documentation License.
%
% See the file doc/licenses/LICENSE for more details.

\section{Introduction}

\beamer\ is a \LaTeX\ class for creating presentations that are held using a projector, but it can also be used to create transparency slides. Preparing presentations with \beamer\ is different from preparing them with \textsc{wysiwyg} programs like OpenOffice.org Impress, Apple Keynote, KOffice KPresenter or Microsoft PowerPoint. A \beamer\ presentation is created like any other \LaTeX\ document: It has a preamble and a body, the body contains |\section|s and |\subsection|s, the different slides (called \emph{frames} in \beamer) are put in environments, they are structured using |itemize| and |enumerate| environments, and so on. The obvious disadvantage of this approach is that you have to know \LaTeX\ in order to use \beamer. The advantage is that if you know \LaTeX, you can use your knowledge of \LaTeX\ also when creating a presentation, not only when writing papers.


\subsection{Main Features}

The list of features supported by \beamer\ is quite long (unfortunately, so is presumably the list of bugs supported by \beamer). The most important features, in our opinion, are:
\begin{itemize}
\item
  You can use \beamer\ with |pdflatex|, |latex|+|dvips|, |lualatex| and |xelatex|. |latex|+|dvipdfm| isn't supported (but we accept patches!).
\item
  The standard commands of \LaTeX\ still work. A |\tableofcontents| will still create a table of contents, |\section| is still used to create structure, and |itemize| still creates a list.
\item
  You can easily create overlays and dynamic effects.
\item
  Themes allow you to change the appearance of your presentation to suit your purposes.
\item
  The themes are designed to be usable in practice, they are not just for show. You will not find such nonsense as a green body text on a picture of a green meadow.
\item
  The layout, the colors, and the fonts used in a presentation can easily be changed globally, but you still also have control over the most minute detail.
\item
  A special style file allows you to use the \LaTeX-source of a presentation directly in other \LaTeX\ classes like |article| or |book|. This makes it easy to create presentations out of lecture notes or lecture notes out of presentations.
\item
  The final output is typically a \textsc{pdf}-file. Viewer applications for this format exist for virtually every platform. When bringing your presentation to a conference on a memory stick, you do not have to worry about which version of the presentation program might be installed there. Also, your presentation is going to look exactly the way it looked on your computer.
\end{itemize}


\subsection{History}

Till Tantau created \beamer\ mainly in his spare time. Many other people have helped by sending him emails containing suggestions for improvement or corrections or patches or whole new themes (by now, this amounts to over a thousand emails concerning \beamer). Indeed, most of the development was only initiated by feature requests and bug reports. Without this feedback, \beamer\ would still be what it was originally intended to be: a small private collection of macros that make using the |seminar| class easier. Till created the first version of \beamer\ for his PhD defense presentation in February 2003. A month later, he put the package on \textsc{ctan} at the request of some colleagues. After that, things somehow got out of hand.

After being unmaintained since 2007, in April 2010 Till handed over the maintenance to Joseph Wright and Vedran Mileti\'c, who are still maintaining it: improving code, fixing bugs, adding new features and helping users.


\subsection{Acknowledgments}

Till Tantau: \emph{``Where to begin? \beamer's development depends not only on me, but on the feedback I get from other people. Many features have been implemented because someone requested them and I thought that these features would be nice to have and reasonably easy to implement. Other people have given valuable feedback on themes, on the user's guide, on features of the class, on the internals of the implementation, on special \LaTeX\ features, and on life in general. A small selection of these people includes (in no particular order and I have surely forgotten to name lots of people who really, really deserve being in this list): Carsten (for everything), Birgit (for being the first person to use \beamer\ besides me), Tux (for his silent criticism), Rolf Niepraschk (for showing me how to program \LaTeX\ correctly), Claudio Beccari (for writing part of the documentation  on font encodings), Thomas Baumann (for the emacs stuff), Stefan M\"uller (for not loosing hope), Uwe Kern (for \textsc{xcolor}), 
Hendri Adriaens (for \textsc{ha-prosper}), Ohura Makoto (for spotting typos). People who have contributed to the themes include Paul Gomme, Manuel Carro, and Marlon R\'egis Schmitz.''}

Joseph Wright: \emph{``Thanks to Till Tantau for the huge development effort in creating \beamer. Sincere thanks to Vedran Mileti\'c for taking the lead in continuing development.''}

Vedran Mileti\'c: \emph{``First, I would like to thank Karl Berry and Sanda Buja\v ci\'c for encouragement, without which I wouldn't ever be anything but a \LaTeX\ user. I would also like to thank Ana Me\v strovi\'c, my colleague, who was excited by the prospect of using \beamer\ for preparing class material; Ivona Frankovi\'c and Marina Rajnovi\'c, my students at Department of Informatics, who were the first to hear about \LaTeX, \beamer\ and how it can help in preparing class material. I would like to thank Heiko Oberdiek (for \textsc{hyperref}), Johannes Braams (for \textsc{babel}) and Philipp Lehman (for \textsc{biblatex}). Above all, I owe a lot to Till Tantau for developing \beamer\ in the first place and to Joseph Wright for developing \textsc{siunitx} and for helping me develop \beamer\ further.''}


\subsection{How to Read this User's Guide}

You should start with the first part. If you have not yet installed the package, please read Section~\ref{section-installation} first. If you are new to \beamer, you should next read the tutorial in Section~\ref{section-tutorial}. When you sit down to create your first real presentation using \beamer, read Section~\ref{section-workflow} where the technical details of a possible workflow are discussed. If you are still new to creating presentations in general, you might find Section~\ref{section-guidelines} helpful, where many guidelines are given on what to do and what not to do. Finally, you should browse through Section~\ref{section-solutions}, where you will find ready-to-use solution templates for creating talks, possibly even in the language you intend to use.

The second part of this user's guide goes into the details of all the commands defined in \beamer, but it also addresses  other technical issues having to do with creating presentations (like how to include graphics or animations).

The third part explains how you can change the appearance of your presentation easily either using themes or by specifying colors or fonts for specific elements of a presentation (like, say, the font used for the numbers in an enumeration).

The fourth part talks about handouts and lecture notes, so called ``support material''. You will frequently have create some kind of support material to give to your audience during the talk or after it, and this part will explain how to do it using the same source that you created your presentation from.

The last part contains ``howtos,'' which are explanations of how to get certain things done using \beamer.

\medskip
\noindent
This user's guide contains descriptions of all ``public'' commands, environments, and concepts defined by the \beamer-class. The following examples show how things are documented. As a general rule, red text is \emph{defined}, green text is \emph{optional}, blue text indicates special mode considerations.

\begingroup
\noindexing
\begin{command}{\somebeamercommand\oarg{optional arguments}\marg{first argument}\marg{second argument}}
  Here you will find the explanation of what the command |\somebeamercommand| does. The green argument(s) is optional. The command of this example takes two parameters.

  \example
  |\somebeamercommand[opt]{my arg}{xxx}|
\end{command}

\begin{environment}{{somebeamerenvironment}\oarg{optional arguments}\marg{first argument}}
  Here you will find the explanation of the effect of the environment |somebeamerenvironment|. As with commands, the green arguments are optional.

  \example
\begin{verbatim}
\begin{somebeamerenvironment}{Argument}
  Some text.
\end{somebeamerenvironment}
\end{verbatim}
\end{environment}

\begin{element}{some beamer element}\yes\yes\yes
  Here you will find an explanation of the template, color, and/or font |some beamer element|. A ``\beamer-element'' is a concept that is explained in more detail in Section~\ref{section-elements}. Roughly speaking, an \emph{element} is a part of a presentation that is potentially typeset in some special way. Examples of elements are frame titles, the author's name, or the footnote sign. For most elements there exists a \emph{template}, see Section~\ref{section-elements} once more, and also a \beamer-color and a \beamer-font.

  For each element, it is indicated whether a template, a \beamer-color, and/or a \beamer-font of the name |some beamer element| exist. Typically, all three exist and are employed together when the element needs to be typeset, that is, when the template is inserted the \beamer-color and -font are installed first. However, sometimes templates do not have a color or font associated with them (like parent templates). Also, there exist \beamer-colors and -fonts that do not have an underlying template.

  Using and changing templates is explained in Section~\ref{section-templates}. Here is the essence: To change a template, you can say
\begin{verbatim}
\setbeamertemplate{some beamer element}{your definition for this template}
\end{verbatim}

  Unfortunately, it is not quite trivial to come up with a good definition for some templates. Fortunately, there are often \emph{predefined options} for a template. These are indicated like this:
  \begin{itemize}
    \itemoption{square}{}
    causes a small square to be used to render the template.
    \itemoption{circle}{\marg{radius}}
    causes circles of the given radius to be used to render the template.
  \end{itemize}

  You can install such a predefined option like this:
\begin{verbatim}
\setbeamertemplate{some beamer element}[square]
%% Now squares are used

\setbeamertemplate{some beamer element}[circle]{3pt}
%% Now a circle is used
\end{verbatim}

  \beamer-colors are explained in Section~\ref{section-colors}. Here is the essence: To change the foreground of the color to, say, red, use
\begin{verbatim}
\setbeamercolor{some beamer element}{fg=red}
\end{verbatim}

  To change the background to, say, black, use:
\begin{verbatim}
\setbeamercolor{some beamer element}{bg=black}
\end{verbatim}

  You can also change them together using |fg=red,bg=black|. The background will not always be ``honoured,'' since it is difficult to show a colored background correctly and an extra effort must be made by the templates (while the foreground color is usually used automatically).

  \beamer-fonts are explained in Section~\ref{section-fonts}. Here is the essence: To change the size of the font to, say, large, use:
\begin{verbatim}
\setbeamerfont{some beamer element}{size=\large}
\end{verbatim}

  In addition to the size, you can use things like |series=\bfseries| to set the series, |shape=\itshape| to change the shape, |family=\sffamily| to change the family, and you can use them in conjunction. Add a star to the command to first ``reset'' the font.
\end{element}

\beamernote
As next to this paragraph, you will sometimes find the word \textsc{presentation} in blue next to some paragraph. This means that the paragraph applies only when you ``normally typeset your presentation using \LaTeX\ or pdf\LaTeX.''

\articlenote
Opposed to this, a paragraph with \textsc{article} next to it describes some behavior that is special for the |article| mode. This special mode is used to create lecture notes out of a presentation (the two can coexist in one file).

\endgroup


\subsection{Getting Help}

When you need help with \beamer, please do the following:
\begin{enumerate}
\item
  Read the user guide, at least the part that has to do with your problem.
\item
  If that does not solve the problem, try searching \href{http://tex.stackexchange.com}{TeX-sx (\texttt{tex.stackexchange.com})}.
  Perhaps someone has already reported a similar problem and someone has found a solution.
\item
  If you find no answers there, or if you are sure you have found a bug in
  \beamer{}, please report it \emph{via} \href{https://github.com/josephwright/beamer/issues}{\texttt{github.com/josephwright/beamer/issues}}.
\item
  Before you file a bug report, especially a bug report concerning the installation, make sure that this is really a bug. In particular, have a look at the |.log| file that results when you \TeX\ your files. This |.log| file should show that all the right files are loaded from the right directories. Nearly all installation problems can be resolved by looking at the |.log| file.

  If you can, before reporting the bug, retest using latest version of \beamer\ with latest version of \TeX\ Live. This can help isolate bugs from other packages that might affect \beamer.
\item
  \emph{As a last resort} you can try emailing authors. We do not mind getting emails, we simply get way too many of them. Because of this, we cannot guarantee that your emails will be answered timely or even at all. Reporting
an issue is usually a better approach as they don't get lost.
\end{enumerate}
