% Copyright 2003--2007 by Till Tantau
% Copyright 2010 by Vedran Mileti\'c
% Copyright 2013,2015 by Vedran Mileti\'c, Joseph Wright
%
% This file may be distributed and/or modified
%
% 1. under the LaTeX Project Public License and/or
% 2. under the GNU Free Documentation License.
%
% See the file doc/licenses/LICENSE for more details.

\section{Solution Templates}
\label{section-solutions}

In the subdirectories of the directory |beamer/solutions| you will find \emph{solution templates} in different languages. A solution template is a \TeX-text that ``solves'' a specific problem. Such a problem might be ``I need to create a 20 minute talk for a conference'' or ``I want to create a slide that introduces the next speaker'' or ``I want to create a table that is uncovered piecewise.'' For such a problem, a solution template consists of a mixture of a template and an example that can be used to solve this particular problem. Just copy the solution template file (or parts of it) and freely adjust them to your needs.

The collecting of \beamer\ solution templates has only begun and currently there are only very few of them. We hope that in the future more solutions will become available and we would like to encourage users of the \beamer\ class to send us solutions they develop. We would also like to encourage users to help in translating solutions to languages other than English and German. If you have written a solution or a translation, please feel free to send it to us (make sure however, that it contains about the same amount of explanations and justifications as do the existing solutions).

The following list of solution templates is sorted by the length of the talks for which they provide a template. As always, the solutions can be found in the directory |beamer/solutions|.

\begin{solution}{short-talks/speaker\_introduction-ornate-2min}
  \begin{itemize}
  \item
    Introducing another speaker.
  \item
    Talk length is about 2min.
  \item
    Ornate style.
  \end{itemize}

  \beamernote
  \TeX-version available in languages |de|, |en|, and |fr|.

\end{solution}

\begin{solution}{generic-talks/generic-ornate-15min-45min}
  \begin{itemize}
  \item
    Generic solution template for talks on any subject.
  \item
    Talk length is between 15min and 45min.
  \item
    Ornate style.
  \end{itemize}

  \beamernote
  \TeX-version available in languages |de|, |en|, and |fr|.

\end{solution}

\begin{solution}{conference-talks/conference-ornate-20min}
  \begin{itemize}
  \item
    Talk at a conference/colloquium.
  \item
    Talk length is about 20 minutes.
  \item
    Ornate style.
  \end{itemize}

  \beamernote
  \TeX-version available in languages |de|, |en|, and |fr|.

\end{solution}
