% Copyright 2003--2007 by Till Tantau
% Copyright 2010 by Vedran Mileti\'c
% Copyright 2012,2015 by Vedran Mileti\'c, Joseph Wright
%
% This file may be distributed and/or modified
%
% 1. under the LaTeX Project Public License and/or
% 2. under the GNU Free Documentation License.
%
% See the file doc/licenses/LICENSE for more details.

\section{Taking Advantage of Multiple Screens}
\label{section-twoscreens}

This section describes options provided by \beamer\ for taking advantage of computers that have more than one video output and can display different outputs on them. For such systems, one video output can be attached to a projector and the main presentation is shown there. The second video output is attached to a small extra monitor (or is just shown on the display of the computer) and shows, for example, special notes for you. Alternatively, the two outputs might be attached to two different projectors. One can then show the main presentation on the first projection and, say, the table of contents on the second. Or the second projection might show a version translated into a different language. Or the second projection might always show the ``previous'' slide. Or \ldots---we are sure you can think of further useful things.

The basic idea behind \beamer's support of two video outputs is the following: Using special options you can ask \beamer\ to create a \pdf-file in which the ``pages'' are unusually wide or high. By default, their height will still be 128mm, but their width will be 192mm (twice the usual default 96mm). These ``superwide'' pages will show the slides of the main presentation on the left and auxiliary material on the right (this can be switched using appropriate options, though hyperlinks will only work if the presentation is on the left and the second screen on the right).

For the presentation you attach two screens to the system. The windowing system believes that the screen is twice as wide as it actually is. Everything the windowing system puts on the left half of this big virtual screen is redirected to the first video output, everything on the right half is redirected to the second video output.

When the presentation program displays the specially prepared superwide \beamer-presentation, exactly the left half of the screen will be filled with the main presentation, the right part is filled with the auxiliary material---voil\`a. Not all presentation programs support this special feature. For example, the Acrobat Reader 6.0.2 will only use one screen in fullscreen mode on MacOS~X. On the other hand, a program named PDF Presenter supports showing dual-screen presentations. Generally, you will have to find out for yourself whether your display program and system support showing superwide presentations stretching over two screens.

\beamer\ uses the package |pgfpages| to typeset two-screen presentations. Because of this, your first step when creating a two-screen presentation is to include this package:
\begin{verbatim}
\documentclass{beamer}
\usepackage{pgfpages}
\end{verbatim}

The next step is to choose an appropriate option for showing something special on the second screen. These options are discussed in the following sections.

One of the things these options do is to setup a certain |pgfpages|-layout that is appropriate for two-screen presentations. However, you can still change the |pgfpages|-layout arbitrarily, afterwards. For example, you might wish to enlarge the virtual pages. For details, see the documentation of |pgfpages|.


\subsection{Showing Notes on the Second Screen}

The first way to use a second screen is to show the presentation on the main screen and to show your notes on the second screen. The option |show notes on second screen| can be used for this. It is described on page~\pageref{command-notesonsecondscreen}.


\subsection{Showing Second Mode Material on the Second Screen}

The second way to use the second screen is to show ``a different version'' of the presentation on the second screen. This different version might be a translation or it might just always be the current table of contents.

To specify what is shown on the second screen, you can use a special \beamer-mode called |second|. This mode behaves similar to modes like |handout| or |beamer|, but its effect depends on the exact options used:

\begin{beameroption}{second mode text on second screen}{|=|\meta{location}}
  This option causes the second screen to show the second mode material. The \meta{location} of the second screen can be |left|, |right|, |bottom|, or |top|.

  In detail, the following happens: When a new frame needs to be typeset, \beamer\ checks whether the special option |typeset second| is given. If not, the frame is typeset normally and the slides are put on the main presentation screen (more precisely, on the logical |pgfpages|-page number zero). The second screen (logical page number one) shows whatever it showed before the frame was typeset.

  If the special frame option |typeset second| is given, after each slide of the frame the frame contents is typeset once more, but this time for the mode |second|. This results in another slide, which is put on the second screen (on logical page number one). Then the whole page is shipped out.

  The |second| mode behaves more like the |beamer| mode than other modes: Any overlay specification for |beamer| will also apply to |second| mode, unless an explicit |second| mode specification is also given. In particular, |\only<1-2>{Text}| will be shown on slides 1 and 2 in |second| mode, but only on the first slide in |handout| mode or |trans| mode.

  \example
\begin{verbatim}
\documentclass{beamer}
\usepackage{pgfpages}
\setbeameroption{second mode text on second screen}
\begin{document}
\begin{frame}[typeset second]
  This text is shown on the left and on the right.
  \only<second>{This text is only shown on the right.}
  \only<second:0>{This text is only shown on the left.}
\end{frame}
\begin{frame}
  This text is shown on the left. The right shows the same as for the
  previous frame.
\end{frame}
\begin{frame}[typeset second]
  \alt<second>{The \string\alt command is useful for second
    mode. Let's show the table of contents, here: \tableofcontents}
  {Here comes some normal text for the first slide.}
\end{frame}
\end{document}
\end{verbatim}

  \example
  The following example shows how translations can be added in a comfortable way.
\begin{verbatim}
\documentclass{beamer}
\usepackage{pgfpages}
\setbeameroption{second mode text on second screen}
\DeclareRobustCommand\translation[1]{\mytranslation#1\relax}
\long\def\mytranslation#1|#2\relax{\alt<second>{#2}{#1}}
\title{\translation{Preparing Presentations|Vortr\"age vorbereiten}}
\author{Till Tantau}
\begin{document}
\begin{frame}[typeset second]
  \titlepage
\end{frame}
\begin{frame}[typeset second]
  \frametitle{\translation{This is the frame title.|Dies ist der Titel des Rahmens.}}
  \begin{itemize}
  \item<1-> \translation{First|Erstens}.
  \item<2-> \translation{Second|Zweitens}.
  \item<3-> \translation{Third|Drittens}.
  \end{itemize}
  \translation{Do not use line-by-line uncovering.|Man sollte Text nicht
  Zeile f\"ur Zeile aufdecken.}
\end{frame}
\end{document}
\end{verbatim}
\end{beameroption}

In the last of the above example, it is a bit bothersome that the option |typeset second| has to be added to each frame. The following option globally sets this option:

\begin{beameroption}{always typeset second mode}{|=|\meta{true or false}}
  When this option is set to true, every following frame will have the option |typeset second| set to true.
\end{beameroption}


\subsection{Showing the Previous Slide on the Second Screen}

\begin{beameroption}{previous slide on second screen}{|=|\meta{location}}
  This option causes the second screen to show the previous slide that was typeset, unless this is overruled by a frame with the |[typeset second]| option set. The idea is that if you have two projectors you can always present ``the last two'' slides simultaneously and talk about them.

  Using this option will switch off the updating of external files like the table of contents.
\end{beameroption}
