% Copyright 2003, 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% This program can be redistributed and/or modified under the terms
% of the GNU Public License, version 2.

\section{Workflow for Creating a Presentation}

\label{section-workflow}

This section presents a possible workflow for creating a \beamer\ 
presentation and possibly a handout to go along with it. Technical
questions are addressed, like which programs to call with 
which parameters, and hints are given on how to create a
presentation. If you have already created numerous presentations, you
may wish to skip the first of the following steps 
and only have a look at how to convert the |.tex| file into a
|.pdf| or |.ps| file.


\subsection{Step Zero: Know the Time Constraints}

When you start to create a presentation, the very first thing you
should worry about is the amount of time you have for your
presentation. Depending on the occasion, this can
be anything between 2 minutes and two hours. A simple rule for the
number of frames is that you should have at most one frame per
minute.

In most situations, you will have less time for your presentation that
you would like. \emph{Do not try to squeeze more into a
  presentation than time allows for.} No matter how important some
detail seems to you, it is better to leave it out, but get the main
message across, than getting neither the main message nor the detail
across. 

In many situations, a quick appraisal of how much time you have will
show that you won't be able to mention certain details. Knowing this can
save you hours of work on preparing slides that you would have to remove
later anyway.




\subsection{Step One: Setup the Files}

\beamernote
It is advisable that you create a folder for each
presentation. Even though your presentation will usually reside in a
single file, \TeX\ produces so many extra files that things can easily
get very confusing otherwise. The folder's name should ideally start
with the date of your talk in ISO format (like 2003-12-25 for a
Christmas talk), followed by some reminder text of what the talk is
all about. Putting the date at the front in this format causes your
presentation folders to be listed nicely when you have several of them
residing in one directory. If you use an extra directory for each
presentation, you can call your main file |main.tex|. 

To create an initial |main.tex| file for your talk, copy an
existing file from the |beamer/solutions| directory and adapt it to
your needs. A list of possible \beamer\ solutions that contain
templates for presentation \TeX-files can be found below.

If you wish your talk to reside in the same file as some different,
non-presentation article version of your text, it is advisable to
setup a more elaborate file scheme. See
Section~\ref{section-article-version-workflow} for details.

\lyxnote
You can either open a new file and then select |beamer| as the
document class or you say ``New from template'' and then use a
template from the directory |beamer/solutions|.


\subsubsection{Solution Templates for Presentations}

The following list of solution templates is sorted by the length of
the talks for which they provide a template. As always, the solutions
can be found in the directory |beamer/solutions|.

\begin{solution}{short-talks/speaker\_introduction-ornate-2min}
  \begin{itemize}
  \item
    Introducing another speaker.
  \item
    Talk length is about 2min.
  \item
    Ornate style.
  \end{itemize}
  \beamernote \TeX-version available in languages |de| and |en|.
  \lyxnote    \LyX-version available in languages |de| and |en|.
\end{solution}

\begin{solution}{generic-talks/generic-ornate-15min-45min}
  \begin{itemize}
  \item
    Generic solution template for talks on any subject.
  \item
    Talk length is between 15min and 45min.
  \item
    Ornate style.
  \end{itemize}
  \beamernote \TeX-version available in languages |de| and |en|.
  \lyxnote    \LyX-version available in languages |de| and |en|.
\end{solution}

\begin{solution}{conference-talks/conference-ornate-20min}
  \begin{itemize}
  \item
    Talk at a conference/colloquium.
  \item
    Talk length is about 20 minutes.
  \item
    Ornate style.
  \end{itemize}
  \beamernote \TeX-version available in languages |de| and |en|.
  \lyxnote    \LyX-version available in languages |de| and |en|.
\end{solution}





\subsection{Step Two: Structure Your Presentation}

With the time constraints in mind, make a mental inventory of the
things you can reasonably talk about within the time available. Then
categorize the inventory into sections and subsections. For very long
talks (like a 90 minute lecture), you might also divide your talk into
independent parts (like a ``review of the previous lecture part'' and
a ``main part''). Put |\section| and |\subsection| commands into
the (more or less empty) main file. Do not create any frames until you
have a first working version of a possible table of contents. Do not
feel afraid to change it later on as you work on the talk.

You should not use more than four sections and not less than two per
part. Even four sections are usually too much, unless they follow 
a very easy pattern. Five and more sections are simply too hard to
remember for the audience. After all, when you present the table of
contents, the audience will not yet really be able to grasp the
importance and relevance of the different sections and will most
likely have forgotten them by the time you reach them.

Ideally, a table of contents should be understandable by itself. In
particular, it should be comprehensible \emph{before} someone has
heard your talk. Keep section and subsection titles
self-explaining. Note that each part has its own table of contents. 

Both the sections and the subsections should follow a logical
pattern. Begin with an explanation of what your talk is all about. (Do
not assume that everyone knows this. The Ignorant Audience Law states:
The audience always knows less than you think it should know, even if
you take the Ignorant Audience Law into account.) Then explain what
you or someone else has found out concerning the subject
matter. Always conclude your talk with a summary that repeats the main
message of the talk in a short and simple way. People pay most
attention at the beginning and at the end of talks. The summary is
your ``second chance'' to get across a message.

You can also add an appendix part using the |\appendix| command. Put
everything into this part that you do not actually intend to talk
about, but that might come in handy when questions are asked.



\subsection{Step Three: Creating a PDF or PostScript File}

\beamernote
Once a first version of the structure is finished, you should create a
first PDF or PostScript file of your (still empty) talk. This file
will only contain the title page and the table of contents. The file
might  look like this:

\begin{verbatim}
\documentclass{beamer}
%% This is the file main.tex

\usepackage{beamerthemesplit}

\title{Example Presentation Created with the Beamer Package}
\author{Till Tantau}
\date{\today}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\section*{Outline}
\frame{\tableofcontents}

\section{Introduction}
\subsection{Overview of the Beamer Class}
\subsection{Overview of Similar Classes}

\section{Usage}
\subsection{...}
\subsection{...}

\section{Examples}
\subsection{...}
\subsection{...}

\frame{} % to enforce entries in the table of contents

\end{document}
\end{verbatim}

The empty frame at the end (which should be deleted later) ensures
that the sections and subsections are actually part of the table of
contents. This frame is necessary since a |\section| or |\subsection|
command following the last page of a document have no effect.

\lyxnote
Use ``View'' to check whether the presentation compiles fine. Note
that you must put the table of contents inside a frame, but that the
title page is created automatically.


\subsubsection{Creating PDF}

\beamernote
To create a |PDF| version of this file, run the program
|pdflatex| on |main.tex| at least twice. Your need to run it twice, so
that \TeX\ can create the table of contents. (It may even be necessary
to run it more often since all sorts of auxiliary files are
created.) In the following example, the greater-than-sign is the prompt. 

\begin{verbatim}
> pdflatex main.tex
    ... lots of output ...
> pdflatex main.tex
    ... lots of output ...
\end{verbatim}

You can next use a program like the Acrobat Reader or |xpdf|
to view the resulting presentation.

\begin{verbatim}
> acroread main.pdf
\end{verbatim}

When printing a presentation using Acrobat, make sure that the option
``expand small pages to paper size'' in the printer dialog is
enabled. This is necessary, because slides are only 128mm times 96mm.

To put several slides onto one page (useful for the handout version)
or to enlarge the slides, you can use the program |pdfnup|. Also, many
commercial programs can perform this task. If you put several slides
on one page and if these slides normally have a white background, it
may be useful to write the following in your preamble:

\begin{verbatim}
\mode<handout>{\beamertemplatesolidbackgroundcolor{black!5}}
\end{verbatim}

This will cause the slides of the handout version to have a very
light gray background. This makes it easy to discern the slides'
border if several slides are put on one page.

\lyxnote
Choose ``View pdf'' to view your presentation.


\subsubsection{Creating PostScript}
\label{section-postscript}

\beamernote
To create a PostScript version, you should first ascertain that the
\textsc{hyperref} package (which is automatically loaded by the
\beamer\ class) uses the option |dvips| or some compatible
option, see the documentation of the \textsc{hyperref} package for
details. Whether this is the case depends on the contents of your
local |hyperref.cfg| file. You can enforce the usage of this
option by passing |dvips| or a compatible option to the
\beamer\ class (write |\documentclass[dvips]{beamer}|), which
will pass this option on to the \textsc{hyperref} package.

You can then run |latex| twice, followed by |dvips|.

\begin{verbatim}
> latex main.tex
    ... lots of output ...
> latex main.tex
    ... lots of output ...
> dvips -P pdf main.dvi
\end{verbatim}

The option (|-P pdf|) tells |dvips| to use
Type~1 outline fonts instead of the usual Type~3 bitmap fonts. You may
wish to omit this option if there is a problem with it. 

If you wish each slide to completely fill a letter-sized page, use the
following commands instead:

\begin{verbatim}
> dvips -P pdf -tletter main.dvi -o main.temp.ps
> psnup -1 -W128mm -H96mm -pletter main.temp.ps main.ps
\end{verbatim}

For A4-sized paper, use:

\begin{verbatim}
> dvips -P pdf -ta4 main.dvi -o main.temp.ps
> psnup -1 -W128mm -H96mm -pa4 main.temp.ps main.ps
\end{verbatim}

In order to create a white margin around the whole page (which is sometimes
useful for printing), add the option |-m 1cm| to the options of
|psnup|. 

To put two or four slides on one page, use |-2|, respectively
|-4| instead of |-1| as the first parameter for
|psnup|. In this case, you may wish to add the option
|-b 1cm| to add a bit of space around the individual slides. The same
trick as for the \pdf-version can be used to make the borders of
slides more pronounced in the handout version.

You can convert a PostScript file to a pdf file using

\begin{verbatim}
> ps2pdf main.ps main.pdf
\end{verbatim}


\lyxnote
Use ``View Postscript'' to view the PostScript version.


\subsection{Step Four: Create Frames}

Once the table of contents looks satisfactory, start creating frames
for your presentation. In the following, some guidelines that I stick
to are given on what to put on slides and what not to put. You can
certainly ignore any of these guidelines, but you should be aware of
it when you ignore a rule and you should be able to justify it to
yourself. 

\lyxnote
To create a frame, use the style ``BeginFrame''. The frame title
is given on the line of this style. The frame ends automatically with
the start of the next frame, with a section or subsection command, and
with an empyt line in the sylte ``EndFrame''. Note that the last frame
of your presentation must be ended using ``EndFrame'' and that the
last frame before the appendix must be ended this way.

\subsubsection{Guidelines on What to Put on a Frame}

\begin{itemize}
\item
  A frame with too little on it is better than a
  frame with too much on it.
\item
  Do not assume that everyone in the audience is an expert on the
  subject matter. (Remember the Ignorant Audience Law.) Even if the
  people listening to you should be experts, they may last have heard
  about things you consider obvious several years ago. You should
  always have the time for a quick reminder of what exactly a
  ``semantical complexity class'' or an ``$\omega$-complete partial
  ordering'' is.
\item
  Never put anything on a slide that you are not going to explain
  during the talk, not even to impress anyone with how
    complicated your subject matter really is. However, you may
  explain things that are not on a slide.
\item
  Keep it simple. Typically, your audience will see a slide for less
  than 50 seconds. They will not have the time to puzzle through long
  sentences or complicated formulas.
\end{itemize}


\subsubsection{Guidelines on Titles}

\begin{itemize}
\item
  Put a title on each frame. The title explains the contents of the
  frame to people who did not follow all details on the slide.
\item
  The title should really \emph{explain} things, not just give a
  cryptic summary that cannot be understood unless one has understood
  the whole slide. For example, a title like ``The Poset'' will have
  everyone puzzled what this slide might be about. Titles like
  ``Review of the Definition of Partially Ordered Sets (Posets)'' or
  ``A Partial Ordering on the Columns of the Genotype Matrix'' are
  \emph{much} more informative.
\item
  Ideally, titles on consecutive frames should ``tell a story'' all by
  themselves.
\item
  In English, you should \emph{either} \emph{always} capitalize all words in
  a frame title except for words like ``a'' or ``the'' (as in a
  title), \emph{or} you \emph{always} use the normal lowercase
  letters. Do \emph{not} mix this; stick to one rule. The same is true
  for block titles. For example, do not use titles like ``A short
  Review of Turing machines.'' Either use ``A Short Review of Turing
  Machines.'' or ``A short review of Turing machines.'' (Turing is
  still spelled with a capital letter since it is a name).
\item
  In English, the title of the whole document should be
  capitalized, regardless of whether you capitalize anything else.
\item
  In German and other languages that have lots of capitalized words,
  always use the correct upper-/lowercase letters. Never capitalize
  anything in addition to what is usually capitalized.
\end{itemize}



\subsubsection{Guidelines on the Body Text}

\begin{itemize}
\item
  \emph{Never} use a smaller font size to ``fit more on a frame.''
\item
  Prefer enumerations and itemize environments over plain text. Do not
  use long sentences.
\item
  Do not hyphenate words. If absolutely necessary, hyphenate words
  ``by hand,'' using the command~|\-|.
\item
  Break lines ``by hand'' using the command~|\\|. Do not rely on
  automatic line breaking. Break where there is a logical pause. For 
  example, good breaks in ``the tape alphabet is larger
  than the input alphabet'' are before ``is'' and before the second
  ``the.'' Bad breaks are before either ``alphabet'' and before
  ``larger.''
\item
  Text and numbers in figures should have the \emph{same} size as
  normal text. Illegible numbers on axes usually ruin a chart and its
  message. 
\end{itemize}


\subsubsection{Guidelines on Graphics}

\begin{itemize}
\item
  Put (at least) one graphic on each slide, whenever
  possible. Visualizations help an audience enormously.
\item
  Usually, place graphics to the left of the text. (Use the
  |columns| environment.) 
\item
  Graphics should have the same typographic parameters as the
  text: Use the same fonts (at the same size) in graphics as in the
  main text. A small dot in a graphic should have exactly the same 
  size as a small dot in a text. The line width should be the same as
  the stroke width used in creating the glyphs of the font. For
  example, an 11pt non-bold Computer Modern font has a stroke width of
  0.4pt.
\item
  While bitmap graphics, like photos, can be much more colorful than the
  rest of the text, vector graphics should follow the same ``color
  logic'' as the main text (like black~= normal lines, red~= hilighted
  parts, green~= examples, blue~= structure).
\item
  Like text, you should explain everything that is shown on a
  graphic. Unexplained details make the audience puzzle whether this
  was something important that they have missed. Be careful when
  importing graphics from a paper or some other source. They usually
  have much more detail than you will be able to explain.
\end{itemize}

For technical hints on how to create graphics, see
Section~\ref{section-graphics}.


\subsubsection{Guidelines on Colors}

\begin{itemize}
\item
  Use colors sparsely. The prepared themes are already quite
  colorful (blue~= structure, red~= alert, green~= example). If you
  add more colors, you should have a \emph{very} good reason.
\item
  Be careful when using bright colors on white background,
  \emph{especially} when using green. What looks good on your monitor
  may look bad during a presentation due to the different ways
  monitors, beamers, and printers reproduce colors. Add lots of black
  to pure colors when you use them on bright backgrounds.
\item
  Maximize contrast. Normal text should be black on white or at least
  something very dark on something very bright. \emph{Never} do things
  like ``light green text on not-so-light green background.''
\item
  Background shadings decrease the legibility without increasing the
  information content. Do not add a background shading just because it
  ``somehow looks nicer.'' In the examples that come along with the
  \beamer\ class, the backgrounds are intended as demonstrations, not
  as recommendations.
\item
  Inverse video (bright text on dark background) can be a problem
  during presentations in bright environments since only a small
  percentage of the presentation area is light up by the
  beamer. Inverse video is harder to reproduce on printouts and on
  transparencies. 
\end{itemize}


\subsubsection{Guidelines on Animations and Special Effects}

\begin{itemize}
\item
  Use animations to explain the dynamics of systems, algorithms, etc.
\item
  Do \emph{not} use animations just to attract the attention of your
  audience. This often distracts attention away from the main topic of the
  slide.
\item
  Do \emph{not} use distracting special effects like ``dissolving''
  slides unless you have a very good reason for using them. If you use
  them, use them sparsely. 
\end{itemize}


\subsubsection{Ways of Improving Compilation Speed}
\label{section-speedup}

While working on your presentation, it may sometimes be useful to
\TeX\ your |.tex| file quickly and have the presentation contain only
the most important information. This is especially true if you have a
slow machine. In this case, you can do several things to speedup the
compilation. First, you can use the |draft| class option.

\begin{classoption}{{draft}}
  Causes the headlines, footlines, and sidebars to be replaced by
  gray rectangles (their sizes are still computed, though). Many
  other packages, including |pgf| and |hyperref|, also ``speedup''
  when this option is given.
\end{classoption}

Second, you can use the following command:

\begin{command}{{\includeonlyframes}\marg{frame label list}}
  This command behaves a little bit like the |\includeonly| command:
  Only the frames mentioned in the list are included. All other frames
  are suppressed. Nevertheless, the section and subsection commands
  are still executed, so that you still have the correct navigation
  bars. By labeling the current frame as, say, |current| and then
  saying |\includeonlyframes{current}|, you can work on a single frame
  quickly.  

  The \meta{frame label list} is a comma-separated list (without
  spaces) of the names of frames that have been labeled. To label a
  frame, you must pass the option |label=|\meta{name} to the |\frame|
  command or |frame| environment.

  \example
\begin{verbatim}
\includeonlyframes{example1,example3}

\frame[label=example1]
{This frame will be included. }

\frame[label=example2]
{This frame will not be included. }

\frame{This frame will not be included.}

\againframe{example1} % Will be included
\end{verbatim}
\end{command}



\subsection{Step Five: Test Your Presentation}

\emph{Always} test your presentation. For this, you should
vocalize or subvocalize your talk in a quiet environment. Typically,
this will show that your talk is too long. You should then remove
parts of the presentation, such that it fits into the allotted time
slot. Do \emph{not} attempt to talk faster in order to squeeze the
talk into the given amount of time. You are almost sure to loose your
audience this way.

Do not try to create the ``perfect'' presentation immediately. Rather,
test and retest the talk and modify it as needed. 




\subsection{Step Six: Optionally Create a Handout or an Article Version}

Once your talk is fixed, you can create a handout, if this seems
appropriate. For this, use the class option |handout| as
explained in Section~\ref{handout}. Typically, you might wish
to put several handout slides on one page. See
Section~\ref{section-postscript} on how to do this.

You may also wish to create an article version of your talk. An
``article version'' of your presentation is a normal \TeX\ text
typeset using, for example, the document class |article| or perhaps
|llncs| or a similar document class. The \beamer\ class offers
facilities to have this version coexist with your presentation version
in one file and to share code. Also, you can include slides of your
presentation as figures in your article version. Details on how to
setup the article version can be found in
Section~\ref{section-article}.  

\lyxnote
Creating an article version is not really possible in \LyX. You
can \emph{try}, but I would not advise it.




